Latent Predicate Networks: Concept Learning with Probabilistic Context-Sensitive Grammars

### 2014-10-20 drafts/cfp/draft.tex
another pass on the draft

{\small For humans, learning abstract concepts and learning language
  go hand in hand: we acquire abstract knowledge primarily through
  linguistic experience, and acquiring abstract concepts is a crucial
  step in learning the meanings of linguistic expressions. How,
  though, can machines learn systems of concepts from sentences about
  those concepts? One approach is to estimate the distribution from
  which those sentences are drawn, and, in doing so, infer the latent
  concepts and relationships that best explain the data. We present
  early work on a learning framework called Latent Predicate Networks
  (LPNs) which learns concepts by inferring the parameters of
  probabilistic context-sensitive grammars over sentences.  We show
  that for a small fragment of sentences expressing relationships
  between English number words, we can use hierarchical Bayesian
  inference to learn LPNs that can answer simple queries about
  previously unseen linguistic concepts within this domain. These
  generalizations demonstrate LPNs' promise as a tool for learning and
  representing conceptual knowledge in language.}

@@

Probabilistic Context Free Grammars (PCFGs) are popular and much
studied generative models for natural language
syntax~\cite{Someone}. While debate continues about their ability to
express the syntax of natural language, it is agreed that they cannot
express dependencies between words that depend upon each other for
their meaning. Consider an example from the domain of sentences
addressed in this paper: ``after twenty eight comes twenty nine.''
The second ``twenty'' depends on the first ``twenty'' and the
``eight'', and the ``nine'' depends on the ``eight''. This pattern is
a clear example of the sort of \emph{cross-serial dependencies} which
context-free grammars cannot support.

@@

\begin{figure}[t]
	\centering
	\begin{minipage}[b]{\linewidth}
1) A(a, a) :: 0.25. \\
2) A(b, b) :: 0.3. \\
3) A(X a, Y a) $\leftarrow$ A(X, Y) :: 0.25. \\
4) A(X b, Y b) $\leftarrow$ A(X, Y) :: 0.2.
		\subcaption{}
		\label{fig:grammar}
	\end{minipage}
	%\rule[0pt]{0.8\linewidth}{.5pt}
	\begin{minipage}[b]{\linewidth}
		\begin{lstlisting}
values('A', [1, 2, 3, 4],
    [0.25, 0.30, 0.25, 0.20]).

reduce('A'-[[a],[a]],1).
reduce('A'-[[b],[b]],2).
reduce('A'-[A2,B2],3) :- lpn('A'-[X,Y]),
    append(X,[a],A2), append(Y,[a],B2).
reduce('A'-[A2,B2],4) :- lpn('A'-[X,Y]),
    append(X,[b],A2), append(Y,[b],B2).

lpn(P-IN) :- msw(P,V), reduce(P-IN,V).
		\end{lstlisting}
		\subcaption{}
		\label{fig:prism}
	\end{minipage}
	\caption{Possible encodings of the 2-copy language as (\subref{fig:grammar}) an LPN, (\subref{fig:prism}) a PRISM program.}
	\label{fig:copy}
\end{figure}

@@

This work is also closely related to the growing field of semantic
parsing. Both are aimed at jointly learning syntactic and semantic
information from a body of sentences. Semantic parsers, however, learn
separate systems of syntax and semantics, often pairing Combinatory
Categorial Grammars of syntax with an explicit logical language of
semantics, often the untyped lambda
calculus~\cite{berant2013semantic,liang2013learning,kwiatkowski2010inducing}. By
contrast, our approach uses a single formalism, the latent predicate
network, to describe both which sentences are syntactically valid and
what sorts of semantic information those sentence imply. Semantic
parsers also typically learn models of semantics by pairing sentences
with an explicit logical form or real-world denotation. In our model
it is the grammar induction itself which reveals candidate relations
between words that are key to capturing their meaning.

More generally, we see this model as helping to strike a balance
between expressivity and tractability, bridge the gap between more
traditional PCFG grammar induction and program induction over more
powerful languages~\cite{dechter2013bootstrap,} (CITATION other
P.I. papers).

>>>>

{\small For humans, learning abstract concepts and learning language
  go hand in hand: we acquire abstract knowledge primarily through
  linguistic experience, and acquiring abstract concepts is a crucial
  step in learning the meanings of linguistic expressions. Number
  knowledge is a case in point: we can only acquire concepts such as
  seventy-three through linguistic means, and we can only know what
  the sentence ``Seventy-three is more than twice as big as thirty-one"
  means if we can grasp the meanings of all its component number
  words. How, though, can machines learn systems of concepts from
  sentences about those concepts? One approach is to estimate the
  distribution from which those sentences are drawn, and, in doing so,
  infer the latent concepts and relationships that best explain the
  data. We present early work on a learning framework called Latent
  Predicate Networks (LPNs) which learns concepts by inferring the
  parameters of probabilistic context-sensitive grammars over
  sentences.  We show that for a small fragment of sentences
  expressing relationships between English number words, we can use
  hierarchical Bayesian inference to learn LPNs that can answer simple
  queries about previously unseen linguistic concepts within this
  domain. These generalizations demonstrate LPNs' promise as a tool
  for learning and representing conceptual knowledge in language.}

@@

Probabilistic Context Free Grammars (PCFGs) are popular and much
studied generative models for natural language
syntax~\cite{Someone}. While debate continues about their ability to
express the syntax of natural language, it is agreed that they cannot
express dependencies between words that depend upon each other for
their meaning. Consider an example from the domain of sentences
addressed in this paper: ``after twenty eight comes twenty nine.''
The second ``twenty'' depends on the first ``twenty'' and the
``eight'', and the ``nine'' depends on the ``eight''. This pattern is
a clear example of the \emph{cross-serial dependencies} which
context-free grammars cannot support.

@@

\begin{figure}[t]
	\centering
	\begin{minipage}[b]{0.8\linewidth}
1) A(a, a) :: 0.25. \\
2) A(b, b) :: 0.3. \\
3) A(X a, Y a) $\leftarrow$ A(X, Y) :: 0.25. \\
4) A(X b, Y b) $\leftarrow$ A(X, Y) :: 0.2.
		\subcaption{}
		\label{fig:grammar}
	\end{minipage}
	%\rule[0pt]{0.8\linewidth}{.5pt}
	\begin{minipage}[b]{0.8\linewidth}
        \fontsize{9}{10.5}\selectfont\ttfamily
		\begin{verbatim}
values(`A', [1, 2, 3, 4],
    [0.25, 0.30, 0.25, 0.20]).

reduce(`A'-[[a],[a]],1).
reduce(`A'-[[b],[b]],2).
reduce(`A'-[A2,B2],3) :- lpn(`A'-[X,Y]),
    append(X,[a],A2), append(Y,[a],B2).
reduce(`A'-[A2,B2],4) :- lpn(`A'-[X,Y]),
    append(X,[b],A2), append(Y,[b],B2).

lpn(P-IN) :- msw(P,V), reduce(P-IN,V).
		\end{verbatim}
		\subcaption{}
		\label{fig:prism}
	\end{minipage}
	\caption{Possible encodings of the 2-copy language as (\subref{fig:grammar}) an LPN, (\subref{fig:prism}) a PRISM program.}
	\label{fig:copy}
\end{figure}

@@

LPNs are only one of many models aimed at jointly learning the
structure and meaning of natural language expressions, with semantic
parsers being a popular alternative. Semantic parsers, however, learn
separate systems of syntax and semantics, often pairing Combinatory
Categorial Grammars of syntax with an explicit logical language of
semantics, typically the untyped lambda
calculus~\cite{berant2013semantic, liang2013learning,
  kwiatkowski2010inducing}. By contrast, our approach uses a single
formalism, the LPN, to describe both which sentences are syntactically
valid and what sorts of semantic information those sentences
imply. Semantic parsers also typically learn models of semantics by
pairing sentences with an explicit logical form or real-world
denotation. In our model it is the grammar induction over sentences
alone which reveals candidate relations between words that are key to
capturing their meaning. Future work should seek to understand the
limits of modeling syntactic and semantic information as inseparable
parts of as a single system, and whether there might be ways to
flexibly combine the strengths of this approach with more traditional
dual system models.

Previous work has shown that generative models of probabilistic
programs are a useful tool for understanding human concept learning
for domains as varied as handwritten characters and structured visual
objects~\cite{dechter2013bootstrap, Stuhlmuller, BrendensPaper}. Here,
we continue that work by showing how some forms of both human language
learning and concept learning may be fruitfully modeled as inference
over a generative model of probabilistic logic programs. Future work
should work explore how LPNs and its extensions may apply to other
domains where generative models of probabilistic programming have
already proven useful, as well as generalizing the current approach to
linguistic concepts.
